Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over convex domains and domains with shellable triangulations

### 2024-10-06 main.tex
last theoretical section

We prepare some further notation for our main result. 
Whenever $\calT$ is an $n$-dimensional triangulation, we use the abbreviation
\begin{align*}
    C_{\PF,\Gamma,k,p}(\calT) 
    :=
    \max\limits_{ \substack{ T \in \calT \\ \dim(T)=n \\ 0 \leq l \leq n } }
    C_{\PF,\Gamma,l,k,p}(T)
    .
\end{align*}
Suppose that $\calT$ is an $n$-dimensional triangulation and has a shelling $T_{0}, T_{1}, \dots, T_{M}$. 
For each $0 \leq m \leq M$, we write 
\begin{align*}
    X_{m} := T_0 \cup T_1 \cup \dots \cup T_{m}.
\end{align*}
For each $1 \leq m \leq M$, the interface of the new simplex to the previous intermediate triangulation is 
\begin{align*}
    \Gamma_{m} := T_{m} \cap X_{m-1} %\left( T_{0} \cup \dots \cup T_{m-1} \right)
    .
\end{align*}
According to Lemma~\ref{lemma:existenceofstar}, 
there exists a simplex $S_{m} \in \calT$ 
such that $T_{m}$ is the last $n$-simplex in the shelling that belongs to $T_{m} \in \patch_{\calT}(S_{m})$.
In particular, already $\patch_{\calT}(S_{m}) \subseteq X_{m}$. % is already included in the union of the simplices up to $T_m$.

@@

The main result of this manuscript for shellable triangulations of domains, not necessarily convex or even star-shaped, is the following theorem.

\begin{theorem}\label{theorem:poincarefriedrichsestimate:exterior}
    Let $\Omega \subseteq \bbR^{n}$ be a domain with a shellable $n$-dimensional triangulation $\calT$.
    Let $T_0, T_1, \dots, T_M$ be a shelling.
    Then for any $u \in W^{p}\Alt^{k}(\Omega)$ 
    there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan w = \cartan u$
	and such that the following estimates hold:
	\begin{gather*}
        \| w_0 \|_{L^{p}(T_{0})} \leq C_{{\PF},k,T_{0},p} \| \cartan u \|_{L^{p}(T_{0})}.
    \end{gather*}
    and for each $T_{m} \in \calT$, $1 \leq m \leq M$ we have the recursive estimate 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        &
        \leq  
        C_{\PF,T_{m},\Gamma_{m},k,p}(\calT) 
        \left( 
            \| \cartan u      \|_{L^{p}(T_{m})} 
            +
            \Cfive{n}{k}(\calT)
            \Cfiveprime{n}{k}(\calT)^{k} 
            \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
            \| \cartan u \|_{L^{p}(\Minako_{m-1})}
        \right)
        \\&\qquad
        + 
        \Cfive{n}{k}(\calT) 
        \Cfiveprime{n}{k}(\calT)^{k-1} 
        \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(\Minako_{m-1})}
        .
    \end{align*}
\end{theorem}
\begin{proof}
    Let $u \in W^{p}\Alt^{k}(\Omega)$. 
    First, there exists $w_0 \in W^{p}\Alt^{k}(T_{0})$ satisfying $\cartan w_0 = \cartan u$ over $T_{0}$ together with 
    \begin{gather*}
        \| w_0 \|_{L^{p}(T_{0})} \leq C_{{\PF},k,T_{0},p} \| \cartan u \|_{L^{p}(T_{0})}.
    \end{gather*}
    %Suppose that claim is true when the domain has $m$ simplices of dimension $n$. 
    Suppose that $0 < m \leq M$ such that there exists $w_{m-1} \in W^{p}\Alt^{k}(X_{m-1})$ 
    with $\cartan w_{m-1} = \cartan u$ over $X_{m-1}$. 
    By assumption, $T_{m}$ and $X_{m-1}$ share the interface $\Gamma_{m}$, which is a collection of faces of $T_{m}$. 
    In accordance to Lemma~\ref{lemma:existenceofstar}, adding $T_{m}$ completes a star in $\calT$ around some simplex $S_{m}$, 
	and we let $\Minako_{m-1} \subseteq X_{m-1}$ be the complement of $T_{m}$ in that newly completed star.

@@

\begin{theorem}\label{theorem:poincarefriedrichsestimate:exterior:contraction}
    Let $\Omega \subseteq \bbR^{n}$ be a domain with an $n$-dimensional shellable triangulation $\calT$.
    Let $T_0, T_1, \dots, T_M$ be a shelling.
    Then for any $u \in W^{p}\Alt^{k}(\Omega)$ 
    there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan w = \cartan u$ 
    such that 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        \leq 
        C_{\PF,T,k,p}(T_{0})
        \prod_{i=1}^{m} 
        C_{???}
        \| \cartan u \|_{L^{p}(\Omega)}
        .
    \end{align*}
\end{theorem}
\begin{proof}
    We prove this via induction. 
    The statement is clearly satisfied when $m=0$ because then we only consider a single simplex. 
    Let us assume that the statement is true for some integer $m-1 \geq 0$.

    We write $X_{m-1} := \cup_{j=0}^{m-1} T_{j}$, 
    which is a triangulated $n$-dimensional submanifold with boundary.
    By assumption, $T_{m}$ and $X_{m-1}$ share the interface $\Gamma_{m} \subset \partial T_{m}$, 
    which is a collection of faces of $T_{m}$. 
    In accordance to Lemma~\ref{lemma:existenceofstar}, adding $T_{m}$ completes a star in $\calT$ around some simplex $S \subseteq T$, 
    and we let $\Minako_{m-1} \subseteq X_{m-1}$ be the complement of $T_{m}$ in that newly completed star.

    We use the bi-Lipschitz mapping $\Xi_{2} : \patch_{\calT}(S) \rightarrow \Minako_{m-1}$ as discussed in Proposition~\ref{proposition:starreflection},
    which acts as the identity along $\partial \Minako_{m-1} \setminus \partial T$.
    By assumption, 
    there exists $w_{m-1} \in W^{p}\Alt^{k}(X_{m-1})$ 
    such that $\cartan w_{m-1} = \Xi_{2}^{-1} \cartan u$ and 
    \begin{align*}
        \| w_{m-1} \|_{L^{p}(T_{m})}
        \leq 
        C_{\PF,T,k,p}(T_{0})
        \prod_{i=1}^{m} 
        C_{???}
        \| \Xi_{2}^{-\ast} \cartan u \|_{L^{p}(\Omega)}
        .
    \end{align*}
    Because pullback and exterior derivative commute, 
    we have $w := \Xi_{2}^{\ast} w_{m-1}$ satisfying $\cartan w = \cartan u$. 
    Moreover, the desired estimate follow from Proposition~\ref{proposition:starreflection}.
    The proof is complete. 
\end{proof}
\mwl{MV: Why a second theorem? Difference of Theorem~\ref{theorem:poincarefriedrichsestimate:exterior:contraction} wrt Theorem~\ref{theorem:poincarefriedrichsestimate:exterior}? Which is better and when? Add more details to the estimates}

\mwl{MV: There is indeed $\diam(T)^{k}$, with the power $k$? Nothing like this appears in Lemma~\ref{lemma:PFexteriorsimplex}. Moreover, so far, all the constants $C_{{\PF},T,p}$, $C_{{\PF},T_1 \cup T_2,p}$ in Lemma~\ref{lemma:poincarefriedrichsoverfacepatch}, $C_{\PF,T,F,p}$ in Lemma~\ref{lemma:mixedbconsimplex}, or $C_{{\PF},k,\Omega,p}$ in Lemma~\ref{lemma:PFexteriorsimplex} had the diameter $\diam(T)$ inside, only the constants in Section~\ref{subsection: PX_convex} have it outside. Finally, $\Omega$ in the indices is missing, whereas $l$ is superfluous. We should unify.}

>>>>

We prepare some further notation for our two main results. 
Whenever $\calT$ is an $n$-dimensional triangulation, we use the abbreviation
\begin{align*}
    C_{\PF,\Gamma,k,p}(\calT) 
    :=
    \max\limits_{ \substack{ T \in \calT \\ \dim(T)=n \\ 0 \leq l \leq n } }
    C_{\PF,\Gamma,l,k,p}(T)
    .
\end{align*}
Suppose that $\calT$ is an $n$-dimensional triangulation and has a shelling $T_{0}, T_{1}, \dots, T_{M}$. 
For each $0 \leq m \leq M$, we write 
\begin{align*}
    X_{m} := T_0 \cup T_1 \cup \dots \cup T_{m},
\end{align*}
which is by Lemma~\ref{lemma:shell_triang_man} a triangulated $n$-dimensional submanifold with boundary. 
For each $1 \leq m \leq M$, the interface of the new simplex to the previous intermediate triangulation is 
\begin{align*}
    \Gamma_{m} := T_{m} \cap X_{m-1} %\left( T_{0} \cup \dots \cup T_{m-1} \right)
    .
\end{align*}
According to Lemma~\ref{lemma:existenceofstar}, 
there exists a simplex $S_{m} \in \calT$ 
such that $T_{m}$ is the last $n$-simplex in the shelling that belongs to $T_{m} \in \patch_{\calT}(S_{m})$.
In particular, already $\patch_{\calT}(S_{m}) \subseteq X_{m}$. % is already included in the union of the simplices up to $T_m$.

@@

The two main results of this manuscript for shellable triangulations of domains, not necessarily convex or even star-shaped, is the following theorems. 
The first one is inspired by the recursive construction of gradient potentials in Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.
The second one is specific to the fact that shellable triangulations are contractible.

\begin{theorem}\label{theorem:poincarefriedrichsestimate:exterior}
    % Let $\Omega \subseteq \bbR^{n}$ be a domain with a shellable $n$-dimensional triangulation $\calT$.
	Let $\calT$ be a shellable $n$-dimensional triangulation, and let the domain $\Omega \subseteq \bbR^{n}$ be the interior of the underlying set of $\calT$.
    Let $T_0, T_1, \dots, T_M$ be a shelling.
    Then for any $u \in W^{p}\Alt^{k}(\Omega)$ 
    there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan w = \cartan u$
	and such that the following estimates hold:
	\begin{gather*}
        \| w_0 \|_{L^{p}(T_{0})} \leq C_{{\PF},k,T_{0},p} \| \cartan u \|_{L^{p}(T_{0})}.
    \end{gather*}
    and for each $T_{m} \in \calT$, $1 \leq m \leq M$ we have the recursive estimate 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        &
        \leq  
        C_{\PF,T_{m},\Gamma_{m},k,p}(\calT) 
        \left( 
            \| \cartan u      \|_{L^{p}(T_{m})} 
            +
            \Cfive{n}{k}(\calT)
            \Cfiveprime{n}{k}(\calT)^{k} 
            \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
            \| \cartan u \|_{L^{p}(\Minako_{m-1})}
        \right)
        \\&\qquad
        + 
        \Cfive{n}{k}(\calT) 
        \Cfiveprime{n}{k}(\calT)^{k-1} 
        \Cdetsechs{n}{k}(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(\Minako_{m-1})}
        .
    \end{align*}
\end{theorem}
\begin{proof}
    Let $u \in W^{p}\Alt^{k}(\Omega)$. 
    First, there exists $w_0 \in W^{p}\Alt^{k}(T_{0})$ satisfying $\cartan w_0 = \cartan u$ over $T_{0}$ together with 
    \begin{gather*}
        \| w_0 \|_{L^{p}(T_{0})} \leq C_{{\PF},k,T_{0},p} \| \cartan u \|_{L^{p}(T_{0})}.
    \end{gather*}
    Suppose that $0 < m \leq M$ such that there exists $w_{m-1} \in W^{p}\Alt^{k}(X_{m-1})$ 
    with $\cartan w_{m-1} = \cartan u$ over $X_{m-1}$. 
    By assumption, $T_{m}$ and $X_{m-1}$ share the interface $\Gamma_{m}$, which is a collection of faces of $T_{m}$. 
    In accordance to Lemma~\ref{lemma:existenceofstar}, adding $T_{m}$ completes a star in $\calT$ around some simplex $S_{m}$, 
	and we let $\Minako_{m-1} \subseteq X_{m-1}$ be the complement of $T_{m}$ in that newly completed star.

@@

\begin{theorem}\label{theorem:poincarefriedrichsestimate:exterior:contraction}
    Let $\Omega \subseteq \bbR^{n}$ be a domain with an $n$-dimensional shellable triangulation $\calT$. % CAUTION 
    Let $T_0, T_1, \dots, T_M$ be a shelling.
    Then for any $u \in W^{p}\Alt^{k}(\Omega)$ 
    there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan w = \cartan u$ 
    such that 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
        \leq 
        \left( 
        \Cacht{n}{k}(\calT)
        \Cachtprime{n}{k}(\calT)^{k} 
        \Cdetsieb{n}{k}(\calT)^{\frac 1 p} 
        \Csieb{n}{k}(\calT)
        \Csiebprime{n}{k}(\calT)^{k-1} 
        \Cdetacht{n}{k}(\calT)^{\frac 1 p} 
        \right)^{M}
        C_{{\PF},k,T_{0},p}
        \| \cartan u \|_{L^{p}(\Omega)}
        .
    \end{align*}
\end{theorem}
\begin{proof}
    We prove this via induction. 
    The statement is clearly satisfied when $M=0$ because then we only consider a single simplex,
    so that the Poincar\'e--Friedrichs inequality over a simplex can be used. 
    Let us assume that the statement is true for some integer $M-1 \geq 0$.

	Recall that $X_M$ is an $n$-dimensional submanifold with boundary with a triangulation that has a shelling $T_0, T_1, \dots, T_{M-1}$. 
	Write $\Omega' \subseteq X_{M}$ for the interior domain of that manifold. 
	By Lemma~\ref{lemma:existenceofstar}, adding $T_{M}$ completes a star in $\calT$ around some simplex $S_{M} \in X_{M-1}$, 
	and we let $\Minako_{M-1} \subseteq X_{M-1}$ be the complement of $T_{M}$ in that newly completed star.

	We introduce the deformation mapping $\Xi_{2} : \patch_{\calT}(S_M) \rightarrow \Minako_{M-1}$ of Proposition~\ref{proposition:starreflection}. 
	This mapping and its inverse are piecewiselinear, and it acts as the identity along $\partial \Minako_{M-1} \setminus \partial T$.
    We extend it to a bi-Lipschitz piecewise linear mapping $\Xi_{2} : X_{M} \rightarrow X_{M-1}$ by requiring that it is the identity over $X_{M-1} \setminus A_{M-1}$.

	Let $u \in W^{p}\Alt^{k}(\Omega)$ and write $g = \cartan u$. We define $g' = \Xi_{2}^{-\ast} \cartan u = \cartan \Xi_{2}^{-\ast} u$. 
    Then Proposition~\ref{proposition:pullbackestimate} and Proposition~\ref{proposition:starreflection} give  
    \begin{align*}
        \| g' \|_{L^{p}(\Omega')}  
        \leq 
        \Cacht{n}{k}(\calT)
        \Cachtprime{n}{k}(\calT)^{k} 
        \Cdetsieb{n}{k}(\calT)^{\frac 1 p} 
        \| g \|_{L^{p}(\Omega)}  
        .
    \end{align*}
    By induction assumption, there exists $w' \in W^{p}\Alt^{k}(\Omega')$ such that $\cartan w' = g'$ and 
    \begin{align*}
        \| w' \|_{L^{p}(\Omega')} 
        \leq 
        \left( 
        \Cacht{n}{k}(\calT)
        \Cachtprime{n}{k}(\calT)^{k} 
        \Cdetsieb{n}{k}(\calT)^{\frac 1 p} 
        \Csieb{n}{k}(\calT)
        \Csiebprime{n}{k}(\calT)^{k-1} 
        \Cdetacht{n}{k}(\calT)^{\frac 1 p} 
        \right)^{M-1}
        C_{{\PF},k,T_{0},p}
        \| g' \|_{L^{p}(\Omega')}  
    \end{align*}
    Finally, $w = \Xi_{2}^{\ast} w' \in W^{p}\Alt^{k}(\Omega)$ satisfies $\cartan w = g$.
    Moreover, Proposition~\ref{proposition:pullbackestimate} and Proposition~\ref{proposition:starreflection} show 
    \begin{align*}
        \| w \|_{L^{p}(\Omega)}  
        \leq 
        \Csieb{n}{k}(\calT)
        \Csiebprime{n}{k}(\calT)^{k-1} 
        \Cdetacht{n}{k}(\calT)^{\frac 1 p} 
        \| w' \|_{L^{p}(\Omega')} 
    \end{align*}
    The induction step is complete, and the desired result follows. 
\end{proof}

\mwl{MV: Why a second theorem? Difference of Theorem~\ref{theorem:poincarefriedrichsestimate:exterior:contraction} wrt Theorem~\ref{theorem:poincarefriedrichsestimate:exterior}? Which is better and when? Add more details to the estimates}
\mwl{MV: There is indeed $\diam(T)^{k}$, with the power $k$? Nothing like this appears in Lemma~\ref{lemma:PFexteriorsimplex}. Moreover, so far, all the constants $C_{{\PF},T,p}$, $C_{{\PF},T_1 \cup T_2,p}$ in Lemma~\ref{lemma:poincarefriedrichsoverfacepatch}, $C_{\PF,T,F,p}$ in Lemma~\ref{lemma:mixedbconsimplex}, or $C_{{\PF},k,\Omega,p}$ in Lemma~\ref{lemma:PFexteriorsimplex} had the diameter $\diam(T)$ inside, only the constants in Section~\ref{subsection: PX_convex} have it outside. Finally, $\Omega$ in the indices is missing, whereas $l$ is superfluous. We should unify.}
